---
output: html_document
editor_options: 
  chunk_output_type: inline
---

# Hands-on R programming

**This second chapter goes through several ways of imputing the data in R, the different notations, how to deal when values are missing and the environments.**

## **R objects**

When talk about R objects we are referring to the data set. The different ways of defining our data set are described below.

###   **Vectors**
Vectors can only save one-dimension data and only one type of data. We have different types of vectors, such as, numerical, text and logical, look at the examples below:

```{r}
v1 <- c(1,2,3,4,5,6) #numerical
v1
```

```{r}
v2 <- c("I", "am", "Idoia") #text
v2
```

```{r}
v3 <- c(TRUE, FALSE, TRUE) #logical
v3
```

Note that R will be able to do mathematical operations with vectors that contain numerical values, but not with the ones that have text. If we want to check is an object is a vector or not, we can write "is.vector" and if we want to know the length of a vector "length" will give it:

```{r}
is.vector(v1)
```

```{r}
length(v1)
```

Keep in mind that a vector can also contain just one value.

####   **Attributes** 
Attributes are pieces of information that we can attach to a vector, we can look at them by "attributes(vector)", where vector should be substituted by the name of the vector that we want to check. For example, if we want to attributed names to the values of the vector "v1" we can write an attribute to them:

```{r}
names(v1)<- c("one", "two", "three", "four", "five", "six") #names if the attribute that we want to add and in brackets the vector that we want to add them to
attributes(v1) #will show us the attributes of the vector "v1"
```

R will display the attributed is we just look at the vector too:

```{r}
v1
```

We can also remove the attributes from a vector:

```{r}
names(v1) <- NULL
```

If now look at the vector, the names of the values won't be there anymore:

```{r}
v1
```

###   **Matrices** 
Matrices are used for storing values in two dimensions. We can for example transform a vector into a matrix by attributing two dimensions:
```{r}
dim(v1) <- c(2,3)
v1
```
R will fill in the matrix column by column by default, if we want to fill it by row we have to set up "byrow = TRUE".
```{r}
m1 <- matrix(v1, nrow =2) #it will fill it by column
m1
```
```{r}
m2 <- matrix(v1, nrow =2, byrow = TRUE) #it will fill it by row
m2
```
###   **List** 
Lists are data structures that can contain elements of different types (such as vectors, matrices, other lists, functions, etc.) in a single object. Note that the elements are not grouped together, and so, it allows us to have vectors of different lengths and characteristics, an example:
```{r}
l1 <- list(10:30, "Idoia", list(TRUE, FALSE))
l1
```

###   **Data frame** 
Data frames are the two dimensional version of a list, they group vectors together into a two-dimensional table, where each vector becomes a column in the table. Therefore, we must have the same type of data in each column, but we can have different types of data between several columns, such as: 
```{r}
df1 <- data.frame(people = c("Victoria", "Thanh", "Idoia"),  
  description = c("student", "boss", "student"), height = c(1.63, 1.70, 1.65))
df1
```
As it can be seen in the previous example, a data frame allows us to give names to the different vectors that we have, in this case "people", "description" and "height"

###   **Arrays** 
Arrays are used for storing values of the same data type in multiple dimensions, for example to sort values into a 3 dimensional cube, like:
```{r}
ar<- array(c(11:14,21:24,31:34), dim = c(2,2,3))
ar
```

### **Playing with the objects**

####   **Class** 
Class refers to the type or category of an object, it will tell us what tyoe if structure do we have. 

####   **Coercion** 