---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Project

**The main objective of this project is to automatize the workflow of the data treatment for the quantification of Polychlorianted alkanes (PCAs).**

## Correct the instrumental with the internal standard (IS):

```{r}
#Load the libraries
library(dplyr)
library(tidyr)
library(tidyverse)

#Clean the environment
rm(list = ls()) 

#Wide the data so each sample is in one row
GC_qToF_CPsFoodResults3 <- GC_qToF_CPsFoodResults |>  
  filter(Quantitative == "YES") |> 
  pivot_wider(id_cols = `Replicate Name`,
              names_from = Molecule, # name of the new column
              values_from = Area #name of the values for the new columns
             ) #data type       

#Make the data in the data frame numeric
pattern <- "^C\\d+" # This pattern matches column names starting with "C" followed by numbers
GC_qToF_CPsFoodResults3 <- mutate_at(GC_qToF_CPsFoodResults3, vars(matches(pattern)), as.numeric) # Apply as.numeric() to columns matching the pattern
GC_qToF_CPsFoodResults3$IS = as.numeric(GC_qToF_CPsFoodResults3$IS)# Define the IS as numeric 

#Divide the area of the homologues by the IS
CorrectedArea <- mutate_at(GC_qToF_CPsFoodResults3, vars(matches(pattern)), 
                           ~ . / GC_qToF_CPsFoodResults3$IS) #It will now do it for all the columns that have the pattern (start the name with "C")

```

## Calculate the recovery

```{r}
# Define the RS as numeric
GC_qToF_CPsFoodResults3$RS = as.numeric(GC_qToF_CPsFoodResults3$RS) 
GC_qToF_CPsFoodResults3$IS = as.numeric(GC_qToF_CPsFoodResults3$IS)

#Filter for the standards that we want to compare to
GC_qToF_CPsFoodResults3|>  select(`Replicate Name`, IS, RS)|> #Select the columns that we need
  filter(str_detect(`Replicate Name`, "Std", negate = FALSE))|> #Select the standards
  mutate(RecoveryStd = IS / RS) |> #Calculate the IS/RS area for the standards 
  summarize(AverageRecovery = mean(RecoveryStd, na.rm = TRUE)) #Calculate the average of the three standrads

#Dividing the IS by RS for the samples
mutate(GC_qToF_CPsFoodResults3,RecoveryArea= IS / RS) |> #Divide IS/RS for all samples
  mutate(Recovery=RecoveryArea/33.7)
print(Recov)
Recovery[Recovery$column_name == some_condition, "column_name"]

STD <- average(Recovery$v)
#The recovery in %
RecoveryPercentage <- mutate_at(Recovery, vars(matches(pattern)), 
                           ~ . *100) 
```

```{r}
Recovery <- mutate(GC_qToF_CPsFoodResults3, GC_qToF_CPsFoodResults3$IS / GC_qToF_CPsFoodResults3$RS)
```



```{r}

###############################################################################
### Calculate sample recovery (using the standard; 1,2,3,4 Tetrachloronaphthalene, TCN)
###############################################################################

# Read in the standard information for calculating the recovery
df_recovery_std <- read.csv(filename_recovery_std, header=TRUE)

# Replace NA Sample_peak_area with 0
df_recovery_std <- df_recovery_std %>% 
  mutate(Std_peak_area_corr=ifelse(is.na(Std_peak_area), 0, Std_peak_area))

# Filter out just the recovery standard congener (TCN) and one other congener (e.g., C10Cl4) to get the internal standard information (all congeners have the same Sample_peak_area_IS)
# Calculate the mass of internal standard (Sample_IS_mass_pg) and recovery standard (Sample_RS_mass_pg) added to samples
df_recovery <- df_melt %>% 
  filter(Congener=="TCN") %>% 
  select(Congener, Sample_name, Sample_peak_area_RS=Sample_peak_area_corr, Sample_RS_vol_uL) %>% 
  left_join(df_melt %>% 
              filter(Congener=="C10Cl4") %>% # Can filter by any congener (the internal standard is the same for all)
              select(Sample_name, Sample_peak_area_IS, Sample_IS_vol_uL)) %>% 
  mutate(Sample_IS_conc_pg_uL=Sample_IS_conc_pg_uL,
         Sample_RS_conc_pg_uL=Sample_RS_conc_pg_uL,
         Sample_IS_mass_pg=Sample_IS_vol_uL*Sample_IS_conc_pg_uL,
         Sample_RS_mass_pg=Sample_RS_vol_uL*Sample_RS_conc_pg_uL)
# Joining with `by = join_by(Sample_name)`

# Add the standard information for calculating recovery
df_recovery$Std_peak_area_RS <- df_recovery_std$Std_peak_area_corr[nrow(df_recovery_std)]
df_recovery$Std_peak_area_IS <- df_recovery_std$Std_peak_area_IS[1]

# Calculate recoveries
df_recovery <- df_recovery %>% 
  mutate(Std_IS_vol_uL=Std_IS_vol_uL,
         Std_RS_vol_uL=Std_RS_vol_uL,
         Std_IS_conc_pg_uL=Std_IS_conc_pg_uL,
         Std_RS_conc_pg_uL=Std_RS_conc_pg_uL,
         Std_IS_mass_pg=Std_IS_vol_uL*Std_IS_conc_pg_uL,
         Std_RS_mass_pg=Std_RS_vol_uL*Std_RS_conc_pg_uL,
         RF_sample=(Sample_peak_area_RS/Sample_RS_mass_pg)/(Sample_peak_area_IS/Sample_IS_mass_pg),
         RF_std=(Std_peak_area_RS/Std_RS_mass_pg)/(Std_peak_area_IS/Std_IS_mass_pg),
         Recovery_pcnt=RF_std/RF_sample*100)

# Merge recovery back in with original sample data
df_melt <- df_melt %>% 
  left_join(df_recovery %>% 
              select(Sample_name, Recovery_pcnt))
# Joining with `by = join_by(Sample_name)`

```






## Prepare the single chain standards

Before quantifying PCAs in R by the deconvolution method, I need to prepare a "calibration" excel to input in the script. In this Excel one calibration curve is built by linear regression for each PCA homologue in each standard

### Build the calibration curves in R studio:

-   

```{r}
#Select the columns that we need
TESTING |>  select(`Replicate Name`, Molecule, `Isotope Label Type`, `Normalized Area`, `Analyte Concentration`) |> filter(`Isotope Label Type`== "Quan") |> #Filter for the Quan
filter(`Molecule` == "C11H19Cl5")|> #Filter for one PCA homologue
filter(str_detect(`Replicate Name`, "11_57_", negate = TRUE)) |> #remove all the other standards 
ggplot(aes(x = `Analyte Concentration`, y = `Normalized Area`)) + #Plot the scater plot
  geom_point() +
  geom_smooth(method = "lm", # uses the linear regression model
              se = FALSE, # this will remove the confidence intervals
              colour = "orange",
              size = 2) +
  theme_classic() +
  ylab("Normalized Area")

##############Normalized area is missing some values in the data frame so, I tried with the Area instead to see if the code was correct

#Slect the columns that we need
TESTING |>  select(`Replicate Name`, Molecule, `Isotope Label Type`, Area, `Analyte Concentration`) |> filter(`Isotope Label Type`== "Quan") |> #Filter for the Quan
filter(`Molecule` == "C11H19Cl5")|> #Filter for one PCA homologue
filter(str_detect(`Replicate Name`, "11_57_", negate = TRUE)) |> #remove all the other standards 
ggplot(aes(x = `Analyte Concentration`, y = Area)) + #Plot the scater plot
  geom_point() +
  geom_smooth(method = "lm", # uses the linear regression model
              se = FALSE, # this will remove the confidence intervals
              colour = "orange",
              size = 2) +
  theme_classic() +
  ylab("Area")



##########################3########Try to generate a loop for the calibration curve, so I don't need to write the homologue and the standards for every calibration curve###########################################################



```


```{r}
# Create an empty character vector
sample_PFOS_info <- character()

# iteratively fill the empty vector using a for loop and paste text with different cells at the i:th position of each selected column
for (i in 1:length(PFAS_insects$ID)) {
  sample_PFOS_info[i] <- paste("This sample: ", 
              PFAS_insects$ID[i], 
              " has PFOS concentration of", 
              round(PFAS_insects$PFOS[i],2 )
```





```{r}
PFOS_water <- as.numeric()
PFOS_sediment <- as.numeric()
PFOS_earthworms <- as.numeric()
PFOS_emerg_aq_ins <- as.numeric()
PFOS_aq_inslarv <- as.numeric()
PFOS_terr_invcons <- as.numeric()

for (i in 1:length(PFAS_insects$Group)){

if(PFAS_insects$Group[i] == "Water"){
  PFOS_water[i] <- PFAS_insects$PFOS[i]
} else if (PFAS_insects$Group[i] == "Sediment") {
  PFOS_sediment[i] <- PFAS_insects$PFOS[i]
} else if (PFAS_insects$Group[i] == "Earthworms") {
  PFOS_earthworms[i] <- PFAS_insects$PFOS[i]
} else if (PFAS_insects$Group[i] == "Emergent aquatic insects") {
  PFOS_emerg_aq_ins[i] <- PFAS_insects$PFOS[i]
} else if (PFAS_insects$Group[i] == "Aquatic insect larvae") {
  PFOS_aq_inslarv[i] <- PFAS_insects$PFOS[i]
} else if (PFAS_insects$Group[i] == "Terrestrial invertebrate consumers") {
  PFOS_terr_invcons[i] <- PFAS_insects$PFOS[i]
}
}
```





















```{r}
#Response factor of each homologue per 1 ng/µL

#For having the same display as in the Excel sheet we first need to transpose our data frame:

t_CorrectedArea <- t(CorrectedArea)
minC9Cl3<- min(t_CorrectedArea["C9H17Cl3", c('V8', 'V9', 'V10', 'V11')])
minC9Cl3 
###########ERROR: it does not find the column names

#Maybe I can fo it without transposing?

minC9Cl3<- min(CorrectedArea[c('9_48_1', '9_48_2', '9_48_3', '9_48_4'), "C9H17Cl3"]) #min for C9H17Cl3 in the standard
minC9Cl3

maxC9Cl3<- max(CorrectedArea[c('9_48_1', '9_48_2', '9_48_3', '9_48_4'), "C9H17Cl3"]) #max for C9H17Cl3 in the standard
maxC9Cl3

####################ERROR: gives NA as the results 

#x and y represent your independent and dependent variables, respectively.
x<- minC9Cl3
y<- maxC9Cl3
# Fit linear regression model
lm_model <- lm(y ~ x) #lm(y ~ x) fits a linear regression model of y on x.

# Extract the slope coefficient
slope <- round(coef(lm_model)[2], digits = 0) #coef(lm_model) retrieves the coefficients of the fitted model, and [2] selects the second coefficient, which corresponds to the slope.



class(CorrectedArea$C9H17Cl3) #Checking if the values are defined as numeric: yes they are :(

filtered_data <- CorrectedArea[complete.cases(CorrectedArea[c('9_48_1', '9_48_2', '9_48_3', '9_48_4'), "C9H17Cl3"]), ]



Min<- min(GC_qToF_CPsFoodResults3$IS)
#=ROUND(SLOPE(INDIRECT(ADDRESS(ROW();COLUMN()-10+MATCH(MIN(N4:U4);N4:U4;0));TRUE):INDIRECT(ADDRESS(ROW();COLUMN()-10+MATCH(MAX(N4:U4);N4:U4;0));TRUE);INDIRECT(ADDRESS(ROW();COLUMN()-18+MATCH(MIN(N4:U4);N4:U4;0));TRUE):INDIRECT(ADDRESS(ROW();COLUMN()-18+MATCH(MAX(N4:U4);N4:U4;0));TRUE));0)
############Note: N4:U4 are the values of the corrected peak are of the standards


#The result from the previous row to this one
#=IF(ISERROR(W4);0;W4)
```
