---
output: html_document
editor_options: 
  chunk_output_type: console
---

# Project

**The main objective of this project is to automatize the workflow of the data treatment for the quantification of Polychlorianted alkanes (PCAs).**

## Correct the instrumental with the internal standard (IS):

```{r}
#Load the libraries
library(dplyr)
library(tidyr)
library(tidyverse)

#Clean the environment
rm(list = ls()) 

#Wide the data so each sample is in one row
GC_qToF_CPsFoodResults3 <- GC_qToF_CPsFoodResults |>  
  filter(Quantitative == "YES") |> 
  pivot_wider(id_cols = `Replicate Name`,
              names_from = Molecule, # name of the new column
              values_from = Area #name of the values for the new columns
             ) #data type       

#Make the data in the data frame numeric
pattern <- "^C\\d+" # This pattern matches column names starting with "C" followed by numbers
GC_qToF_CPsFoodResults3 <- mutate_at(GC_qToF_CPsFoodResults3, vars(matches(pattern)), as.numeric) # Apply as.numeric() to columns matching the pattern
GC_qToF_CPsFoodResults3$IS = as.numeric(GC_qToF_CPsFoodResults3$IS)# Define the IS as numeric 

#Divide the area of the homologues by the IS
CorrectedArea <- mutate_at(GC_qToF_CPsFoodResults3, vars(matches(pattern)), 
                           ~ . / GC_qToF_CPsFoodResults3$IS) #It will now do it for all the columns that have the pattern (start the name with "C")

```

## Calculate the recovery

```{r}
# Define the RS as numeric
GC_qToF_CPsFoodResults3$RS = as.numeric(GC_qToF_CPsFoodResults3$RS) 

#Dividing the IS by RS
Recovery <- mutate(GC_qToF_CPsFoodResults3, GC_qToF_CPsFoodResults3$IS / GC_qToF_CPsFoodResults3$RS) 

Recovery[Recovery$column_name == some_condition, "column_name"]

STD <- average(Recovery$v)
#The recovery in %
RecoveryPercentage <- mutate_at(Recovery, vars(matches(pattern)), 
                           ~ . *100) 
```

```{r}
Recovery <- mutate(GC_qToF_CPsFoodResults3, GC_qToF_CPsFoodResults3$IS / GC_qToF_CPsFoodResults3$RS)
```

## Prepare the single chain standards

```{r}
#Response factor of each homologue per 1 ng/µL

#For having the same display as in the Excel sheet we first need to transpose our data frame:

t_CorrectedArea <- t(CorrectedArea)
minC9Cl3<- min(t_CorrectedArea["C9H17Cl3", c('V8', 'V9', 'V10', 'V11')])
minC9Cl3 
###########ERROR: it does not find the column names

#Maybe I can fo it without transposing?

minC9Cl3<- min(CorrectedArea[c('9_48_1', '9_48_2', '9_48_3', '9_48_4'), "C9H17Cl3"]) #min for C9H17Cl3 in the standard
minC9Cl3

maxC9Cl3<- max(CorrectedArea[c('9_48_1', '9_48_2', '9_48_3', '9_48_4'), "C9H17Cl3"]) #max for C9H17Cl3 in the standard
maxC9Cl3

####################ERROR: gives NA as the results 

#x and y represent your independent and dependent variables, respectively.
x<- minC9Cl3
y<- maxC9Cl3
# Fit linear regression model
lm_model <- lm(y ~ x) #lm(y ~ x) fits a linear regression model of y on x.

# Extract the slope coefficient
slope <- round(coef(lm_model)[2], digits = 0) #coef(lm_model) retrieves the coefficients of the fitted model, and [2] selects the second coefficient, which corresponds to the slope.



class(CorrectedArea$C9H17Cl3) #Checking if the values are defined as numeric: yes they are :(

filtered_data <- CorrectedArea[complete.cases(CorrectedArea[c('9_48_1', '9_48_2', '9_48_3', '9_48_4'), "C9H17Cl3"]), ]



Min<- min(GC_qToF_CPsFoodResults3$IS)
#=ROUND(SLOPE(INDIRECT(ADDRESS(ROW();COLUMN()-10+MATCH(MIN(N4:U4);N4:U4;0));TRUE):INDIRECT(ADDRESS(ROW();COLUMN()-10+MATCH(MAX(N4:U4);N4:U4;0));TRUE);INDIRECT(ADDRESS(ROW();COLUMN()-18+MATCH(MIN(N4:U4);N4:U4;0));TRUE):INDIRECT(ADDRESS(ROW();COLUMN()-18+MATCH(MAX(N4:U4);N4:U4;0));TRUE));0)
############Note: N4:U4 are the values of the corrected peak are of the standards


#The result from the previous row to this one
#=IF(ISERROR(W4);0;W4)
```

Comparative between a linear calibration in Skyline, R studio and excel:

- 
```{r}
#Slect the columns that we need
TESTING |>  select(`Replicate Name`, Molecule, `Isotope Label Type`, `Normalized Area`, `Analyte Concentration`) |> filter(`Isotope Label Type`== "Quan") |> #Filter for the Quan
filter(`Molecule` == "C11H19Cl5")|> #Filter for one PCA homologue
filter(str_detect(`Replicate Name`, "11_57_", negate = TRUE)) |> #remove all the other standards 
ggplot(aes(x = `Analyte Concentration`, y = `Normalized Area`)) + #Plot the scater plot
  geom_point() +
  geom_smooth(method = "lm", # uses the linear regression model
              se = FALSE, # this will remove the confidence intervals
              colour = "orange",
              size = 2) +
  theme_classic() +
  ylab("Normalized Area")


#Slect the columns that we need
TESTING |>  select(`Replicate Name`, Molecule, `Isotope Label Type`, `Normalized Area`, `Analyte Concentration`) |> filter(`Isotope Label Type`== "Quan") |> #Filter for the Quan
filter(`Molecule` == "C11H19Cl5")|> #Filter for one PCA homologue
filter(str_detect(`Replicate Name`, "11_57_", negate = TRUE)) |> #remove all the other standards 
ggplot(aes(x = `Analyte Concentration`, y = `Normalized Area`)) + #Plot the scater plot
  geom_point() +
  geom_smooth(method = "lm", # uses the linear regression model
              se = FALSE, # this will remove the confidence intervals
              colour = "orange",
              size = 2) +
  theme_classic() +
  ylab("Normalized Area")


```
```
```{r}
PFAS_insects_long <- Statistics_Kumla_20200408 %>% # shortcut for pipe Ctrl + shift + M
  pivot_longer(cols = PFBA:PFOA, # select all columns from PFBA to sum_21PFAS
               names_to = "somePFAS", # name of the new column
               values_to = "Conc") %>%  #name of the values from PFBA:sum_21PFAS
  mutate(somePFAS = as.factor(somePFAS)) #make PFAS column a factor variable
```

```{r}
#check the PFAS group names
PFAS_insects_long |> group_by(somePFAS) |> tally() |> view() 
# select PFOS isomers
PFOS_branched <- c("1-PFOS 419/169", "3/4/5-PFOS80/98.9")

PFAS_insects_long |>
  filter(Group != "Water") |> #remove all water samples to only include solid samples
  filter(str_detect(somePFAS, "sum_", negate = TRUE)) |> #remove all sum of PFAS using the argument: negate=TRUE 
ggplot(aes(x = somePFAS, y = Conc)) +
  geom_point() +
  geom_boxplot() +
  theme_classic() +
  ylab("Concentration of PFAS (ng/g) in solid samples")
  
